\documentclass[smallextended]{svjour3}
\smartqed

% Packages
\usepackage[T1]{fontenc}            % 8-bit font encoding
\usepackage[utf8]{inputenc}         % Should be default, but let's be safe
\usepackage{microtype}              % Subliminal refinements towards typographical perfection
\usepackage[scaled=.8]{beramono}    % Nice-looking monospace font
\usepackage[abbreviations]{foreign} % \eg, \ie, etc.
\usepackage[all]{nowidow}
\usepackage[
	colorlinks=true,
	allcolors=black
]{hyperref}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage[normalem]{ulem}       % Provides \sout
\usepackage{todonotes}
\usepackage{mathptmx}
\usepackage{comment}
\usepackage{subcaption}
%\usepackage{cleveref}               % The almighty \Cref command; keep me last

\journalname{Empirical Software Engineering}

\begin{document}

\title{What Are They Doing? A Commit-Level Study of Hyperactive AI-Augmented Developers on GitHub}

\author{Collective}

\institute{
}

\maketitle

%-----------------------------------------------------------------------
\begin{abstract}
AI coding agents have enabled a new class of developers who commit code at volumes an order of magnitude above historical norms, yet the commit-level behavior of these hyperactive AI-augmented developers remains unstudied.
We present the first longitudinal, commit-level study of such developers on GitHub, based on a curated dataset of seven individuals observed over six months (September 2025 to February 2026): 76,749 commits across 260 repositories, with per-commit file-level detail.
We address four research questions covering temporal activity patterns, coding agent usage and co-evolution, repository-switching behavior, and commit quality.
Our key findings are: (1) hyperactive developers follow divergent trajectories, ranging from sustained monotonic growth to sharp burst-collapse patterns, revealing that high volume is not synonymous with sustained engagement; (2) agent attribution is detectable in 47.6\% of commits, with \texttt{claude\_code} accounting for the majority (36,493 commits), yet two of our seven developers show near-zero hard agent signals despite extreme commit volumes, exposing a critical gap in current detection heuristics; (3) AI agents enable fluid, within-day repository switching at a scale unseen in pre-agent studies; and (4) commit quality signals diverge substantially between agent-attributed and non-attributed commits.
\end{abstract}

\keywords{AI-augmented development \and coding agents \and GitHub mining \and commit analysis \and empirical software engineering}

%-----------------------------------------------------------------------
\section{Introduction}
\label{sec:intro}

% Context
\begin{itemize}
  \item AI coding agents (Claude Code, Cursor, Copilot, Codex, \ldots) entered mainstream use in 2024--2025.
  \item Some developers reportedly reach commit volumes an order of magnitude above historical norms.
  \item Prior large-scale studies (e.g., Robbes et al.~\cite{Robbes2026}) measure adoption rates across many projects but do not zoom in on individual hyperactive developers.
\end{itemize}

% Problem / gap
\begin{itemize}
  \item \emph{What does the commit-level behavior of a hyperactive AI-augmented developer actually look like?}
  \item Temporal rhythm, repository diversification, agent evolution, and commit quality are all unstudied at this granularity.
\end{itemize}

% Contribution
\begin{itemize}
  \item We present a curated dataset of seven hyperactive AI-augmented developers observed over six months (Sep 2025 -- Feb 2026): 76,749 commits, 260 repositories.
  \item We define and operationalize four research questions (RQ1--RQ4) and run quantitative analyses on the dataset.
  \item Dataset, scripts, and figures are publicly available at \url{https://github.com/tdegueul/what-are-they-doing}.
\end{itemize}

% Paper roadmap
\begin{itemize}
  \item Section~\ref{sec:related}: related work.
  \item Section~\ref{sec:dataset}: dataset construction and methodology.
  \item Section~\ref{sec:rqs}: research questions.
  \item Section~\ref{sec:results}: results.
  \item Section~\ref{sec:discussion}: discussion and threats.
  \item Section~\ref{sec:conclusion}: conclusion.
\end{itemize}

%-----------------------------------------------------------------------
\section{Related Work}
\label{sec:related}

% Agent adoption at scale
\begin{itemize}
  \item Robbes et al.~\cite{Robbes2026}: first large-scale study of coding-agent adoption on GitHub (129k projects); 15.85\%--22.60\% adoption; agent-assisted commits are larger and more feature/bugfix-oriented; introduces the \texttt{agent-mining} heuristics library used in our study.
  \item Xu et al.~\cite{Xu2026}: measurable LLM-induced shifts in coding style (snake\_case, complexity, maintainability) across 20k+ repositories linked to arXiv papers.
\end{itemize}

% LLM-generated code and self-admission
\begin{itemize}
  \item Yu et al.~\cite{Yu2024}: self-admitted LLM-generated code in 229 GitHub projects; 696 snippets across five languages; characterises provenance signals.
\end{itemize}

% AI agent PRs
\begin{itemize}
  \item Alam et al.~\cite{Alam2026}: analysis of 8,106 fix-related PRs from five AI coding agents; investigates why agent-involved PRs remain unmerged.
\end{itemize}

% Gap addressed by this work
\begin{itemize}
  \item No prior work studies the \emph{individual} developer trajectory at commit resolution over months; our case-study approach is complementary to population-level studies.
\end{itemize}

%-----------------------------------------------------------------------
\section{Dataset Construction}
\label{sec:dataset}

\subsection{Developer Identification and Selection}
\begin{itemize}
  \item Candidate sources: X/Twitter AI-coding conversations, GitHub commit-volume queries, Gitista, committers.top, claudecount.com leaderboard.
  \item Four-step inclusion workflow: (1) initial screening by activity signals, (2) heuristic agent detection via \texttt{agent-mining}, (3) manual repository and commit review, (4) expert judgment.
  \item Deliberately \emph{curated} rather than mechanically filtered; explicit agent traces strengthen evidence but are not required.
  \item Final cohort: seven developers (Table~\ref{tab:devs}); false positives documented in the repository.
\end{itemize}

\subsection{Data Collection}
\begin{itemize}
  \item GitHub GraphQL API; monthly per-developer JSON files (\texttt{data/\{developer\}-\{YYYY\}-\{MM\}.json}).
  \item Observation window: 2025-09-01 to 2026-02-28 (six months).
  \item Per-commit file-level details cached in \texttt{data/commits/\{sha\}.json}.
  \item 76,749 total commits; 260 tracked repositories; peak aggregate day 2026-02-16 (1,833 commits).
\end{itemize}

\subsection{Agent Detection}
\begin{itemize}
  \item Heuristics from \texttt{agent-mining}~\cite{Robbes2026}: co-author trailers (e.g., \texttt{Co-authored-by: Claude}), commit-message patterns, agent-specific configuration file traces.
  \item 47.6\% of commits carry at least one hard agent signal; eight distinct agent labels detected; \texttt{claude\_code} dominates (36,493 commits).
\end{itemize}

\subsection{Developer Profiles}
% Short characterisation of each developer to motivate the case studies.
\begin{itemize}
  \item \textbf{Dicklesworthstone}: 42,549 commits; 100 repos; 77.1\% agent commits; builds agentic tooling with \textasciitilde20 Claude Max subscriptions.
  \item \textbf{steipete}: 21,222 commits; 71 repos; creator of ClawBot/OpenClaw; highly visible on social media.
  \item \textbf{teamchong}: 5,799 commits; 10 repos; burst-and-collapse trajectory peaking Dec 2025.
  \item \textbf{ruvnet}: 5,660 commits; 9 repos; 84.5\% agent commits; consultant building high-impact software.
  \item \textbf{obra}: 2,656 commits; 36 repos; 52.3\% agent commits; using AI to build a new company.
  \item \textbf{mavam}: 2,543 commits; 21 repos; CEO working on security data pipelines.
  \item \textbf{philipp-spiess}: 801 commits; 15 repos; open-source contributor with a late adoption spike.
\end{itemize}

\begin{table}[t]
\caption{Developer coverage summary (Sep 2025 -- Feb 2026).}
\label{tab:devs}
% TODO: populate from summarize-dataset.py output after data analysis runs.
\centering
\begin{tabular}{lrrrrr}
\toprule
Developer & Repos & Commits & Active days & Agent \% & Top agent \\
\midrule
Dicklesworthstone & 100 & 42,549 & 127 & 77.1 & claude\_code \\
steipete          &  71 & 21,222 & 130 &  0.3 & codex \\
teamchong         &  10 &  5,799 & 108 &  0.6 & claude\_code \\
ruvnet            &   9 &  5,660 & 117 & 84.5 & claude\_code \\
obra              &  36 &  2,656 & 116 & 52.3 & claude\_code \\
mavam             &  21 &  2,543 & 154 & 38.6 & claude\_code \\
philipp-spiess    &  15 &    801 &  69 &  0.2 & opencode \\
\bottomrule
\end{tabular}
\end{table}


\section{What is an average developer behavior on Github pre 2024?} \label{sec:baseline}
%\todo{Romain & Thomas}
could be merged with RW
We only reuse published literature on this topic, for instance:

Oh, and another thing we can add to the paper is referencing this GitClear study to put that productivity in context:
excellent idea. as more academic sources:
- The empirical commit frequency distribution of open source projects
- Free and Open Source Software organizations: A large-scale analysis of code, comments, and commits frequency
Agentic meta-analysis gives "medians vary widely (from 6 to 555 commits per year) depending on whether the dataset includes all repositories or filters for active/popular projects." https://asta.allen.ai/share/185c0e62-9094-480f-897c-7dface083212
- Also, ``\textit{TGIF: the evolution of developer commit times}''~\cite{Karyotakis2026}
- ``\textit{The sky is not the limit: multitasking across GitHub projects}''~\cite{Vasilescu2016sky} (general context on typical commit volumes of prolific developers pre-agents, baseline for RQ4: 37\% of developer-weeks involve contributing to multiple repos, 95\% of devs didn't contribute to more than 1.5 projects each day on average, but 98\% of devs contribute to multiple projects a day at least once. They also interviewed some devs that stated they'd like to increase the number of projects they contribute to.)


%-----------------------------------------------------------------------
\section{Research Questions}
\label{sec:rqs}

\subsection*{RQ1 — Temporal Activity Patterns (@Shane and team)}
\emph{How do commit volumes and activity rhythms evolve over the observation window?}
% acceleration rate, burst events, intra-day patterns
% commit velocity
\begin{itemize}
  \item Motivation: why is this RQ important?
  \item Metrics: daily/monthly commit counts, active days, peak-day counts, intra-day UTC histograms.
  \item Expected analyses: \texttt{plot-commits-over-time-separate.py}, \texttt{histogram-commit-time.py}.
  \item Hypothesis: divergent trajectories (monotonic growth, plateau, burst-collapse) reflect different adoption maturity levels.
\end{itemize}

As a first step towards understanding the characteristics of hyperactive AI-augmented developers, we set out to study their temporal working patterns.
To do so, we plot the development activity of each sampled developer across the studied timeframe with the goal of identifying patterns that may generalize.
We operationalize development activity in terms of (a) the number of commits that were produced; and (b) the churn introduced by those commits.

% TODO: Explain the velocity and acceleration measurements.
In order to find emerging working patterns, we study development velocity,
defined as the rate of development activity over time within the studied
timeframe with a sliding window of seven days. Development velocity indicates
for each AI-augmented developers the intensity of their commits contributions.
Furthermore, we study their development acceleration rate, computed as the
change in development velocity. Development acceleration rate points out whether
developers' activity are increasing, decreasing or stabilizing over time.

\input{rq2.tex}

\subsection*{RQ3 — Commit Quality (@Thomas/Romain and team)}
\emph{What is the quality of commits produced by hyperactive AI-augmented developers?}

\begin{itemize}
  \item Motivation: Just like human developers, agents can introduce bugs, vulnerabilities, and regressions in codebases. Massive amounts of AI-written code raises the question of code review and to which extent human developers can retain oversight over AI-generated code. This RQ investigates the quality of commits (size, focus) and the quality of the underlying code (readability, maintainability).
  \item Metrics:
  \begin{itemize}
  	\item Simple commit-level metrics (files touched, scattering, patch size)
  	\item Simple language-agnostic code quality metrics: function size, LoC, CC, comments, tests
  	\item ``dirty'' commit rate; (@Martin: wdym?)
  	\item Some RW rely on ASAT (\eg SonarQube~\cite{he2025speed}), but that's probably intractable on our dataset, and probably provides little value.
  \end{itemize}
  \item Expected analyses: \texttt{analyze-commit-quality.py}, \texttt{analyze-commits.py}, \texttt{visualize-file-touches.py}
  \begin{itemize}
  	\item Compare agent-authored vs not-agent-authored commits on these metrics
  	\item We can compare across repositories / developers / agents => what has the most influence? Developer's workflows/tooling vs. agents/LLMs vs. repository type/domain.
  \end{itemize}
  \item Hypothesis: high-volume developers produce more scattered (multi-file, multi-language) commits; agent-attributed commits score higher on structural quality signals.
\end{itemize}

\subsection*{RQ4 — Repository Switching Behavior (@Martin and team)}
\emph{How do hyperactive AI-augmented developers switch across repositories?}

Prior work suggests that developers seek to multitask across projects, but that excessive multitasking may negatively affect productivity~\cite{Vasilescu2016sky}. While most developers contribute to multiple repositories at least occasionally, sustained day-to-day multitasking across repositories remains comparatively uncommon. 
%
AI coding agents may alter these dynamics by autonomously executing implementation tasks after instruction, allowing developers to work on other repositories while the agent continues operating. Combined with higher-level prompting interfaces that require less continuous repository-specific interaction, this may reduce the practical costs of switching between repositories and enable more fluid multi-repository workflows.
%
We hypothesize that reduced context-switching costs in AI-agent-assisted workflows are associated with higher within-day repository diversity, and lower concentration relative to pre-agent baselines (cf. Section~\ref{sec:baseline}). 

A repository switch is defined as two consecutive commits targeting different repositories. 
%
We analyze repository-switching behavior using three complementary measures: 
(i) the number of actively maintained repositories per developer-week, defined as repositories receiving at least two commits within the week based on prior work~\cite{Damota2022deep}; higher values may indicate broader concurrent activity enabled by lower context-switching costs; %make it a week
(ii) repository concentration measured using the Gini coefficient, where values near 1 indicate activity concentrated in a small number of repositories and lower values indicate more evenly distributed work; and %measured over a week 
(iii) repository switch counts at the day and session levels, where unusually high values may reflect highly fragmented or strongly parallelized development workflows. We segment activity into sessions using a 30-minute inactivity threshold, per prior work on developer activity sessionization~\cite{Mozannar2024when}. 
%
Together, these measures allow us to assess whether AI-agent-assisted workflows are associated with broader and more fluid multi-repository development behavior relative to pre-agent baselines.

%-----------------------------------------------------------------------
\section{Results}
\label{sec:results}

% All subsections are stubs; detailed findings will be filled in after data analysis scripts are run.

\subsection{RQ1: Temporal Activity Patterns}
\begin{itemize}
  \item TODO: aggregate growth figure and key trajectory types (monotonic, plateau, burst-collapse).
  \item TODO: intra-day UTC distribution per developer (work-hour vs.\ around-the-clock patterns).
  \item TODO: quantify burst events (days with $>100$ commits).
\end{itemize}

\subsection{RQ2: Coding Agent Usage and Co-evolution}
\begin{itemize}
  \item TODO: per-developer agent signal rates over time (line charts from \texttt{plot-agent-commit-coevolution.py}).
  \item TODO: agent diversity index; share of commits with $>1$ agent signal.
  \item TODO: comparison of agent-attributed vs.\ non-attributed commit characteristics.
\end{itemize}


\subsection{RQ3: Commit Quality and Characteristics}
\begin{itemize}
  \item TODO: distribution of ``good'' vs.\ ``dirty'' commit scores across developers.
  \item TODO: Conventional Commits type breakdown (feat, fix, chore, \ldots) per developer.
  \item TODO: language hotspots and test/docs coupling ratios.
  \item TODO: median diff size and file-touch count stratified by agent signal presence.
\end{itemize}

\subsection{RQ4: Repository Switching Behavior}

- Active repositories per week: everyone: mean = 7.20 active repos / developer-week (95\% bootstrap CI: [5.94, 8.69]).
  - Per-developer (min, max, mean, median active repos/week):
    - Dicklesworthstone — 0, 93, 28.15, 26.5
    - steipete          — 1, 55, 11.28, 7.5
    - teamchong         — 0, 8, 2.45, 2.0
    - mavam             — 1, 15, 5.70, 5.0
    - steveyegge        — 0, 3, 1.78, 2.0
    - obra              — 0, 11, 4.24, 4.0
  - some developers (Dicklesworthstone, steipete) maintain activity across a large number of repositories each week, while others (teamchong, steveyegge) show activity on fewer repositories. 

- Gini (concentration): overall median Gini across developer-weeks = 0.449. However, there is significant variation across developers, with some showing much higher concentration (e.g., steipete and Dicklesworthstone) and others showing more even distribution (e.g., teamchong and steveyegge). Looking also at the active repositories per week, we can see that the developers with higher Gini (Dicklesworthstone, steipete) also have higher active repo counts, which shows that while they are actively working on multiple repositories, they still have a few dominant ones. In contrast, developers with lower Gini (teamchong, steveyegge) have fewer active repos per week, suggesting more focused activity. Overall, these findings indicate that AI agents indeed may enable some developers to maintain activity across a wider range of repositories, in line with our hypothesis. 
  - Selected per-developer median Gini: Dicklesworthstone 0.686; steipete 0.748; teamchong 0.350; mavam 0.483; steveyegge 0.092; obra 0.488.

- Repository switches (day / session): Most developers only work on a single repository during most sessions, but switch repositories between sessions, with most of them switching multiple times a day. All developers show some bursty behavior where they switch repositories frequently during certain periods, but they do not maintain this behavior consistently. This suggests that while AI agents may enable more fluid switching, the typical workflow still involves focused work on one repository at a time. However, there are some extreme outliers in terms of switching behavior, with some developers showing very high maxima of repository switches in a day over a longer period of time (e.g., Dicklesworthstone, steipete), which may indicate highly parallelized workflows. 
  - Day-level maxima: Dicklesworthstone = 1292, steipete = 355, clubanderson = 194, teamchong = 154.
  - Day-level medians: Dicklesworthstone = 168.5, steipete = 18.0, teamchong = 2.0.
  - Session-level medians: Dicklesworthstone = 4.0; most other developers = 0.0.
  - Day-level median across developers: 4.0 switches/day.
  - Session-level median across developers: 0.0 switches/session.

- Notable patterns: a small number of developers (Dicklesworthstone, steipete) drive high breadth and extreme daily switch maxima; most developers show low median daily and session switches.

- Caveat: some high-volume days were sampled (see `sampled\_commits` vs `reported\_commits` in `data/exports-RQ4`), so absolute maxima may be affected by collection choices.

- Comparison to \cite{Vasilescu2016sky}: For week-level data, they find developers contributing to a mean of 3.7 repositories, with a minimum of 2, median of 3, and max of 18. Our observed mean of 7.20 active repos/week is higher, with a much higher max (93), suggesting that AI agents may indeed enable broader multi-repository activity. The median of 5.0 active repos/week is also higher than their median of 3, indicating that the increase in breadth is not solely driven by outliers.


\subsection{External Verification through Interviews}

\begin{itemize}
  \item In order to verify and cross-check our results, we additionally invited each developer to participate in a semi-structured interview. The semi-structured approach was chosen for its balance of flexibility and consistency, allowing us to answer our research questions while adapting the interview to each developer's unique workflow and project contexts.
  \item In order to ensure scientific rigor and minimize bias, we developed an interview guide based on our research questions, our domain-specific knowledge, and our general insights from the analysis, as outlined by \cite{Kallio2016}.
  \item After confirming that a semi-structured interview was an appropriate research method for our purposes, we started with our research questions, and then used our domain knowledge to generate specific main themes, top-level questions, and follow up questions. Afterwards we ran an internal pilot test to iterate on our preliminary guide and check for issues with intelligibility, relevance, coverage, and biases. The resulting guide is available in our repository for transparency and replication purposes.
  \item TODO insights from interviews
\end{itemize}

%-----------------------------------------------------------------------
\section{Discussion}
\label{sec:discussion}

\subsection{Threats to Validity}
\begin{itemize}
  \item \emph{Construct}: agent detection relies on co-author trailers and message patterns; developers who suppress signals (steipete) are undercounted, interviews aren't guaranteed to reveal actual developer behavior.
  \item \emph{Internal}: selection bias --- curated cohort; not a random sample; findings may not generalise. changes in frequency or behavior could be driven by external factors
  \item \emph{External}: seven developers, one six-month window; early-adoption period; behaviors may differ post-2026.
  \item \emph{Reliability}: GitHub API rate limits forced sampling on high-volume developers; daily totals may undercount.
\end{itemize}

%-----------------------------------------------------------------------
\section{Conclusion}
\label{sec:conclusion}

\begin{itemize}
  \item We present the first commit-level longitudinal study of hyperactive AI-augmented developers on GitHub.
  \item Key finding (TODO: fill in after analysis): \ldots
  \item The curated dataset and analysis scripts are released to enable replication and extension.
  \item Future work: expand cohort (candidate developers under evaluation), extend observation window, compare against matched non-AI developer baselines.
\end{itemize}

%-----------------------------------------------------------------------
\bibliographystyle{spbasic}
\bibliography{references}

\end{document}
