\subsection*{RQ2: Coding Agent Harness Usage (@Mira and team)}
\emph{Which coding agent harnesses do hyperactive developers use, and for which tasks?}


% Note, use agent harness (or just harness) to clarify we mean the software around the agent, not individual agent instances or the model used by the agent. See fabis original comment below.
% (Fabi): should we consistently refer to it as "agent harnesses" to make it obvious that we're talking about different types of agents () as opposed to just multiple agent instances or sub-agents (i.e. multiple terminals all running an instance of Claude Code)?


% \begin{itemize}
%   \item Motivation; why is this RQ important?
%   \item Metrics: agent signal rate per month, agent diversity per developer, agent-switch events.
%   \item Expected analyses: \texttt{detect-agents.py}, \texttt{plot-agent-commit-coevolution.py}.
%   \item Hypothesis: agent concentration increases over time; \texttt{claude\_code} dominates late in the window; developers switch primary agents.
%   \item
% \end{itemize}


To study coding agent harness usage, we analyze repository activity at the commit level and infer which harness was used for each commit. We define a \emph{coding agent harness} as the software interface through which a developer interacts with an LLM-based coding agent, such as Claude Code, OpenAI Codex, Cursor, or Pi. We define an \emph{agent signal} as observable evidence linking a commit to a specific harness. A commit is assigned to a harness if at least one of the following conditions holds:

\begin{itemize}
\item the commit author name or co-author trailers mention a known harness;
\item the commit modifies harness-specific configuration or metadata files;
\item the commit message contains explicit harness identifiers.
\end{itemize}

% documentation on commit trailers:
% amp has them, yet we basically never see them (95 commits total)
% https://ampcode.com/manual#:~:text=Enable%20adding%20Amp%2DThread%20trailer%20in%20git%20commits%2E%20When%20disabled%2C%20commits%20made%20by%20the%20agent%20will%20not%20include%20the%20Amp%2DThread%3A%20%3Cthread%2Durl%3E%20trailer

A commit without such evidence is classified as \emph{unattributed}. Some harnesses append attribution metadata by default (e.g., Claude Code or Codex co-author trailers), while others leave little or no commit-level trace. Consequently, the absence of a detectable signal does not imply the absence of agent usage. Since this research question requires a harness assignment for all commits, unattributed commits are assigned using a fallback strategy. For each repository and developer, we determine the dominant harness among attributed commits within the surrounding development period and assign unattributed commits to that harness. We evaluate the sensitivity of this approximation during manual validation. We compute the following metrics:

\begin{itemize}
\item \textbf{Agent signal rate}: proportion of commits per month containing detectable harness signals.
\item \textbf{Harness frequency}: number and proportion of commits attributed to each harness.
\item \textbf{Harness diversity}: number of distinct harnesses used by a developer within a time window.
\item \textbf{Harness-switch events}: changes in the dominant harness used by a developer or repository over time.
\item \textbf{Cross-repository reuse}: extent to which the same harnesses appear across multiple repositories owned by the same developer.
\end{itemize}

\paragraph{Task classification.}
To analyze which tasks developers use agents for, we categorize commits using the Conventional Commits specification\footnote{\url{https://www.conventionalcommits.org/en/v1.0.0/}}. Prior work by Robbes et al.~\cite{Robbes2026} found that agent-assisted commits frequently follow these conventions, making them suitable for commit-level task categorization. For each harness, we analyze the distribution of commit categories (e.g., \texttt{feat}, \texttt{fix}, \texttt{refactor}, \texttt{docs}) to identify differences in harness usage across development tasks.


\begin{comment}
  This RQ is about understanding which coding agent harnesses hyperactive developers use and for which tasks. We want to see if there are patterns in the types of agents used, how they evolve over time, and how they relate to the tasks being performed.

To understand how these developers work, we need to understand \emph{which} agent harnesses they use and how they use them.
We hypothesize that hyperactive developers use different coding agent harnesses, such as Claude Code, OpenAI Codex, or Pi, depending on the task, and may change their primary harness over time.

% RM: Discussion point: The original RQ mentioned that we believe that harnesses concentrate on claude code. I dont see how this can be argued, claude code kinda was first, so whatever use we see is a diversification, not concentration?
\todo{add an identifier per hypothesis, such that you can refer to them in the result secton}Our hypothesis is that agent usage increases over time both in frequency and volume, that developers sometimes switch their primary harness, that Claude Code is the most popular harness, that different harnesses are used for different tasks, and that the same harnesses are used across repositories.
\todo{this is a result, it's impossible to have this in the methodology section}We note that the harnesses themselves evolved rapidly during our observation window (e.g., Cursor expanded from completions to agentic workflows), so observed changes may reflect tool availability as much as developer behavior.
To test this, we analyze each commit's metadata and changed files for signals that indicate the use of a specific harness. We assume that all code is generated with some harness, but that not every commit carries a detectable signal. We do note that there are instances of what appear to be fully-manual commits\todo{what's the oracle for that?} but those are entirely drowned out by generated ones, so assuming the use of a harness for all commits is a reasonable approximation of the actual behavior.

We assign a harness to a commit if the commit mentions a harness in its author list or co-author trailers, or if the commit changes agent-specific configuration files.
Notably, attribution signals are uneven: harnesses like Claude Code and Codex append co-author trailers by default,\todo{do we have a primary source to backup this claimi wold be great (but I agree this is almost "common knowledge"} while others (e.g., Copilot, Cursor) leave no commit-level trace, so the absence of a signal does not imply the absence of a harness.

To gain even more insight into what they use agents for, we also analyze the type of work done in the commit based on the categories in the Conventional Commits specification\footnote{\url{https://www.conventionalcommits.org/en/v1.0.0/}}.
Robbes et al.~\cite{Robbes2026} found that agent-assisted commits tend to adhere to these conventions well, which allows us to use them as a reliable categorization.
This reveals which tasks developers use agents for and how they distribute different harnesses across their workflow.

Since our goal here is to study what harnesses are used for, we require a harness assignment for every commit. When no signal is found, we assign the commit to a repository- or developer-specific fallback harness. To determine per-repository default harnesses, we extend the analysis to additional areas of interest, such as pull requests, code reviews, and issue triage, all of which can also carry agent signals. 

\end{comment}

% TODO, this is more conclusion/results text.
% In order to verify our analysis and claims, we took two approaches: (1) we manually reviewed relevant GitHub activity for each developer to collect a qualitative understanding of their workflow and agent usage patterns, and sampled specific commits, commit messages, commit patterns, pull requests, and other activity to check for consistency with our automated analysis; (2) we contacted each of the seven developers for further comment, including a list of questions. These questions were designed to allow us to either confirm or reject our results, based on the developers' answers.

